State of the Practice for Medical Imaging Software

### 2022-03-22 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Revisions to Introduction section, updated RQs

@@

We present a general method to assess the state of the practice for Scientific
Computing (SC) software and apply this method to Medical Imaging (MI) software.
We selected 29 MI software projects from 48 candidates, assessed 10 software
qualities (installability, correctness/ verifiability, reliability, robustness,
usability, maintainability, reusability, understandability,
visibility/transparency and reproducibility) by answering 103 questions for each
software project, and interviewed 8 of the 29 development teams. Based on the
quantitative data for the first 9 qualities, we ranked the MI software with the
Analytic Hierarchy Process (AHP). The top three software products were
\textit{3D Slicer}, \textit{ImageJ}, and \textit{OHIF Viewer}. By interviewing
the developers, we identified three major pain points: i) lack of resources; ii)
difficulty balancing between compatibility, maintainability, performance, and
security; and, iii) lack of access to real-world datasets for testing. For
future MI software projects, we propose adopting test-driven development, using
continuous integration and continuous delivery (CI/CD), using git and GitHub,
maintaining good documentation, supporting third-party plugins or extensions,
considering web application solutions, and improving collaboration between
different MI software projects.

@@

\begin{keyword}
	medical imaging, scientific computing, software engineering, software
	quality, Analytic Hierarchy Process, developer interviews
\end{keyword}

@@

We aim to study the state of software development practice for Medical Imaging
(MI) software.  MI tools use images of the interior of the body (from sources
such as Magnetic Resonance Imaging (MRI), Computed Tomography (CT), Positron
Emission Tomography (PET) and Ultrasound) to provide information for diagnostic,
analytic, and medical applications \citep{FDA2021, enwiki:1034887445,
Zhang2008}.  Given the importance of MI software and the high number of
competing software projects, we wish to understand the merits and drawbacks of
the current development processes, tools and methodologies.  We aim to assess
through a software engineering lens the quality of the existing software with
the hope of highlighting standout examples, understanding current pain points
and providing guidelines and recommendations for future development.

\subsection{Purpose} \label{sec_motivation}

Not only do we wish to gain insight into the state of the practice for MI
software, we also wish to understand the development of research software in
general. We wish to understand the impact of the often cited gap, or chasm,
between software engineering and scientific programming \citep{Storer2017}.
Although scientists spend a substantial proportion of their working hours on
software development \citep{Hannay2009, Prabhu2011}, many developers learn
software engineering skills by themselves or from their peers, instead of from
proper training \citep{Hannay2009}. \citet{Hannay2009} observe that many
scientists showed ignorance and indifference to standard software engineering
concepts. For instance, according to a survey by \citet{Prabhu2011}, more than
half of their 114 subjects did not use any proper debugger for their software.

To gain insights, we devised five research questions, which can be applied to
MI, as well as to other domains, of research software \citep{SmithEtAl2021}. The
questions are designed to learn about the community's interest in, and
experience with, artifacts, tools, principles, processes, methodologies and
qualities.  When we mention artifacts we mean documents, scripts and code that
constitutes a software development project. Example artifacts include
requirements, specifications, user manuals, unit tests, system tests, usability
tests, build scripts, API (Application Programming Interface) documentation,
READMEs, license documents, process documents, and code.  We aimed to collect
enough information to understand the current pain points experienced by the MI
software development community so that we can make some preliminary
recommendations for future improvements.

\begin{description}

\item[RQ1.] \hypertarget{rq1}What artifacts are present in current MI software
projects? What role does documentation play in the projects? What are the
developers' attitude toward it?
\item[RQ2.] \hypertarget{rq2}What tools are used in the development of MI packages?
\item[RQ3.] \hypertarget{rq3}What principles, processes, and methodologies are
used in the development of current MI packages?
\item[RQ4.] \hypertarget{rq4}What are the pain points for developers working on
MI software projects? What aspects of the existing processes, methodologies, and
tools do they consider as potentially needing improvement? What changes to
processes, methodologies, and tools can improve software development and
software quality?
\item[RQ5.] \hypertarget{rq5}What is the current status of the following
software qualities for LBM projects? What actions have the developers taken to
address them?  (Section~\ref{ch_background} provides definitions of the software
qualities.)
\begin{itemize}
	\item Installability
	\item Correctness \& Verifiability
	\item Reliability
	\item Robustness
	\item Usability
	\item Maintainability
	\item Reusability
	\item Understandability
	\item Visibility/Transparency
	\item Reproducibility
\end{itemize}
\item[RQ6.] \hypertarget{rq6}How does software designated as high quality by
this methodology compare with top-rated software by the community?
\end{description}

@@

% Based on our literature survey, we divided MI software into five sub-groups and
% several sub-sub-groups by their major functions, as shown in
% Figure~\ref{fig_mi_functions}.

% \begin{figure}[!ht]
% \centering
% \includegraphics[scale=0.5]{figures/FunctionsOfMiSoftware.pdf}
% \caption{Major functions of MI Software} \label{fig_mi_functions}
% \end{figure}

% \begin{figure}[!ht]
% \centering
% \begin{tikzpicture}[mindmap, grow cyclic, every node/.style=concept, concept
% color=orange!40,
% level 1/.append style={level distance=5cm,sibling angle=72},
% level 2/.append style={level distance=3cm,sibling angle=90},
% level 3/.style={level distance=2.3cm,sibling angle=90}]
% \node{MI software}
% child[concept color=teal!40] { node {Enhancement}}
% child[concept color=teal!40] { node {Analysis}
%     child[concept color=blue!30] { node {Registration}}
%     child[concept color=blue!30,] { node {Segmentation}}
%     child[concept color=blue!30] { node {Statistical Analysis}
%     child[concept color=blue!20] { node {Feature Extraction}}
%     child[concept color=blue!20] { node {Classification}}
%     child[concept color=blue!20] { node {Interpretation}}
%     }}
% child[concept color=teal!40] { node {Simulation/\\Modeling}}
% child[concept color=teal!40] { node {Supporting}
%     child[concept color=blue!30] { node {Tool Kit}}
%     child[concept color=blue!30] { node {PACS}}
% }
% child[concept color=teal!40] { node {Visualization}
%     child[concept color=blue!30] { node {2D Display}}
%     child[concept color=blue!30] { node {3D Rendering}}
%     child[concept color=blue!30] { node {Reformatted Views}}
% };
% \end{tikzpicture}
% \caption{Major functions of MI software}
% \label{fig_mi_functions}
% \end{figure}

@@

We designed a general methodology to assess the state of the practice for SC
software \citep{SmithEtAl2021}. Details can be found in Section~\ref{ch_methods}
and in \citep{Dong2021}.  This methodology builds off prior work to assess the
state of the practice for such domains as Geographic Information Systems
\citep{smith2018state}, Mesh Generators \citep{smith2016state}, Seismology
software \citep{Smith2018Seismology}, and Statistical software for psychology
\citep{smith2018statistical}.  In keeping with the previous methodology, we have
maintained the constraint that the work load for measuring a given domain should
be feasible for a team as small as one person, and for a short time, ideally
around a person month of effort. A person month is considered to be $20$ working
days ($4$ weeks in a month, with $5$ days of work per week) at $8$ person hours
per day, or $20 \times 8 = 160$ person hours.

With our methodology, we first choose an SC domain (in the current case MI) and
identify a list of about 30 software packages. (For measuring MI we used 29
software packages.)  We approximately measure the qualities of each package by
filling in a grading template. Compared with our previous methodology, the new
methodology also includes repository based metrics, such as the number of files,
number of lines of code, percentage of issues that are closed, etc.  With the
quantitative data in the grading template, we rank the software with the
Analytic Hierarchy Process (AHP) (Details are found in
Section~\ref{ch_background}). After this, as another addition to our previous
methodology, we interview some of the development teams to further understand
the status of their development process. Finally, we summarize the measurements
(Section~\ref{ch_results}), summarize the interview answers
(Section~\ref{ch_interview}), answer the research questions
(Section~\ref{ch_answers}) and propose recommendations
(Section~\ref{ch_recommendations}) for future SC software development.

@@

In this section, we introduce the relevant software categories (Section
\ref{sec_software_categories}). The categories will be used when collecting and
assessing the software packages.  We also review the definitions of software
qualities (Section \ref{sec_software_quality}) and provide an overview of the
AHP (Section \ref{sec_AHP}).

@@

Quality is defined as a measure of the excellence or worth of an entity.  As is
common practice, we do not think of quality as a single measure, but rather as a
set of measures.  That is, quality is a collection of different qualities, often
called ``ilities.''  Below we list the 10 qualities of interest for this study,
as summarized in \citet{SmithEtAl2021-QDef}.  The order of the qualities follows
the order used in \cite{GhezziEtAl2003}, which puts related qualities (like
correctness and reliability) together.  Moreover, the order is roughly the same
as the order qualities are considered in practice.  \wss{To be completed - not
currently in sync with QDef paper, QDef paper not completely done.}

@@

If there are $n$ alternatives, for criterion $i = 1, 2, ... , m$, we can create
an $n\times n$ matrix $B_i$ to record the relative preferences between these
choices for each criterion. The way of generating $B_i$ is similar to the one
for $A$. However, rather than comparing the importance between criteria, we
pairwise decide how much we favor one alternative over the other. We use the
same method to calculate the local priority vector for each $B_i$.  The local
priority vector in this case ranks the $n$ alternatives for criterion $i$.

@@

Domain experts may be recruited from academia or industry.  The only
requirements are knowledge of the domain and a willingness to be engaged in the
assessment process.  The Domain Expert does not have to be a software developer,
but they should be a user of domain software.  Given that the domain experts are
likely to be busy people, the measurement process cannot put to much of a burden
on their time.  For the current assessment, our Domain Expert (and paper
co-author) is Dr.\ Michael Noseworthy, Professor of Electrical and Computer
Engineering at McMaster University, Co-Director of the McMaster School of
Biomedical Engineering, and Director of Medical Imaging Physics and Engineering
at St. Joseph’s Healthcare, Hamilton, Ontario, Canada.

@@

The repository measures help us in wo ways. Firstly, they help us with get a
fast and accurate project overview. For example, the number of commits over the
last 12 months shows how active a project has been, and the number of stars and
forks may reveal its popularity. Secondly, the results may affect our decisions
regarding the grading scores for some software qualities. For example, if the
percentage of comment lines is low, we double-check the
understandability of the code; if the ratio of open versus closed pull
requests is high, we pay more attention to maintainability.

@@

@@

We identified the use of unit testing in less than half of the 29 projects. On
the other hand, the interviewees believed that testing (including usability
tests with users) was the top solution to improve correctness, usability, and
reproducibility. One pain point in the development process is the lack of access
to real-world datasets for testing. The developers' strategies to address this
are summarized in Section~\ref{sec_pain_points_3} \ad{to be fixed}. One threat to correctness is:
with huge datasets for testing, the tests are expensive and time-consuming.
Three interviewees endorsed self tests / automated tests, which may save time
for testing.

@@

We answer this question by grading the qualities of the software, collecting
ratings from the MI community, then conducting two comparisons:
\begin{itemize}
\item comparing our ranking with the community ratings on GitHub, such as GitHub
stars, number of forks, and number of people watching the projects (Section~\ref{sec_grading_software});
\item comparing top-rated software designated by our methodology with the ones
recommended by our domain experts (Section \ref{sec_software_selection}).
\end{itemize}

\wss{In the shorter journal paper version, the focus could be just on answering the 
research questions and deleting the previous two sections.}

@@

\item \textbf{Remove privacy from sensitive data.} For data with sensitive
information, a team can ask the data owner to remove such information or add
noise to protect privacy. One example is using deidentified copies of medical
images for testing.

@@

% Overview of scientific computing
We define Scientific Computing (SC) as ``the use of computer tools to analyze or
simulate mathematical models of real world systems of engineering or scientific
importance so that we can better understand and predict the system’s behaviour''
\citep{Smith2006}. Many researchers consider SC as the third pillar of science
and engineering, along with theory and experiment \citep{Landau2005}. Almost all
areas in science and engineering use computers for modeling \citep{Golub2014},
and software plays an essential role in modern scientific research
\citep{Hannay2009, Wilson2014}.  Software development in SC depends on three
fields of knowledge: engineering or scientific domain knowledge, mathematical
algorithm knowledge, and computational algorithm knowledge \citep{Landau2005,
Mehta2015}. Thus, most SC software developers are scientists in SC domains
\citep{Wilson2014}. However, they do not always use modern software development
techniques, tools, and methods \citep{Wilson2014}. Therefore, we developed a
methodology for assessing the state of the practice for SC software
\citep{SmithEtAl2021}. We apply this process to Medical Imaging (MI) software
that belongs to a specific domain of SC.

>>>>

\newcounter{rqnum} %research question number
\newcommand{\rqtherqnum}{RQ`'\therqnum}
\newcommand{\rqref}[1]{RQ\ref{#1}}

\newcounter{pnum} %pain point number
\newcommand{\ppthepnum}{P`'\thepnum}
\newcommand{\ppref}[1]{P\ref{#1}}

@@

We present the state of the practice for Medical Imaging software. We selected
29 medical imaging projects from 48 candidates, assessed 10 software qualities
(installability, correctness/ verifiability, reliability, robustness, usability,
maintainability, reusability, understandability, visibility/transparency and
reproducibility) by answering 108 questions for each software project, and
interviewed 8 of the 29 development teams. Based on the quantitative data for
the first 9 qualities, we ranked the MI software with the Analytic Hierarchy
Process (AHP). The top three software products were \textit{3D Slicer},
\textit{ImageJ}, and \textit{OHIF Viewer}. By interviewing the developers, we
identified three major pain points: i) lack of resources; ii) difficulty
balancing between compatibility, maintainability, performance, and security;
and, iii) lack of access to real-world datasets for testing. For future MI
software projects, we propose adopting test-driven development, using continuous
integration and continuous delivery (CI/CD), using git and GitHub, maintaining
good documentation, supporting third-party plugins or extensions, considering
web application solutions, and improving collaboration between different MI
software projects. \wss{Update after the paper has been revised.}

@@

\begin{keyword}
	medical imaging, research software, software engineering, software
	quality, Analytic Hierarchy Process, developer interviews
\end{keyword}

@@

We aim to study the state of software development practice for Medical Imaging
(MI) software.  MI tools use images of the interior of the body (from sources
such as Magnetic Resonance Imaging (MRI), Computed Tomography (CT), Positron
Emission Tomography (PET) and Ultrasound) to provide information for diagnostic,
analytic, and medical applications \citep{FDA2021, enwiki:1034887445,
Zhang2008}.  An example medical image is provided in Figure~\ref{Fig_Example}.
Given the importance of MI software and the high number of competing software
projects, we wish to understand the merits and drawbacks of the current
development processes, tools and methodologies.  We aim to assess through a
software engineering lens the quality of the existing software with the hope of
highlighting standout examples, understanding current pain points and providing
guidelines and recommendations for future development.

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[scale=0.42]{figures/Example_Aorta.png}        
    \end{center}
    \caption{Example medical image showing the segmentation of an aorta}
    \label{Fig_Example}
\end{figure}

\subsection{Research Questions} \label{sec_motivation}

Not only do we wish to gain insight into the state of the practice for MI
software, we also wish to understand the development of research software in
general. We wish to understand the impact of the often cited gap, or chasm,
between software engineering and scientific programming \citep{Kelly2007,
Storer2017}. Although scientists spend a substantial proportion of their working
hours on software development \citep{Hannay2009, Prabhu2011}, many developers
learn software engineering skills by themselves or from their peers, instead of
from proper training \citep{Hannay2009}. \citet{Hannay2009} observe that many
scientists showed ignorance and indifference to standard software engineering
concepts. For instance, according to a survey by \citet{Prabhu2011}, more than
half of their 114 subjects did not use a proper debugger when coding.

To gain insights, we devised 10 research questions, which can be applied to MI,
as well as to other domains, of research software \citep{SmithEtAl2021}.  The
questions are designed to learn about the community's interest in, and
experience with, artifacts, tools, principles, processes, methodologies and
qualities.  When we mention artifacts we mean the documents, scripts and code
that constitutes a software development project. Example artifacts include
requirements, specifications, user manuals, unit tests, system tests, usability
tests, build scripts, API (Application Programming Interface) documentation,
READMEs, license documents, process documents, and code.  Once we have learned
what MI developers do, we then put this information in context by contrasting MI
software against the trends shown by developers in other research software
communities.  Our aim is to collect enough information to understand the current
pain points experienced by the MI software development community so that we can
make some preliminary recommendations for future improvements.

The questions are used to structure the discussion in the paper, so for each
research question below we point to the section that contains our answer.  We
start with identifying the relevant examples of MI software:

\begin{enumerate}
	\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
	software projects exist, with the constraint that the source code must be
	available for all identified projects? (Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Which
	of the projects identified in \rqref{RQ_WhatProjects} follow current best
	practices, based on evidence found by experimenting with the software and
	searching the artifacts available in each project's repository?
	(Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
	similar is the list of top projects identified in \rqref{RQ_HighestQuality}
	to the most popular projects, as viewed by the scientific community?
	(Section~\ref{Sec_VsCommunityRanking})
    \item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
	do MI projects compare to research software in general with respect to the
	artifacts present in their repositories?
	(Section~\ref{Sec_CompareArtifacts})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:]
	How do MI projects compare to research software in general with respect to
	the use of tools (Section~\ref{Sec_CompareTools}) for:
	\begin{enumerate} 
		\item [\rqref{RQ_CompareToolsProjMngmnt}.a] development; and,
		\item [\rqref{RQ_CompareToolsProjMngmnt}.b] project management?
	\end{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
	How do MI projects compare to research software in general with respect to
	principles, processes and methodologies used?
	(Section~\ref{Sec_CompareMethodologies})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on MI software projects?
	(Section~\ref{ch_interview})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from MI compare to the pain points
	for research software in general? (Section~\ref{ch_interview})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
	developers what specific best practices are taken to address the pain points
	and software quality concerns? (Section~\ref{ch_interview})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practice could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}).
	(Section~\ref{ch_recommendations})

@@

@@

We designed a general methodology to assess the state of the practice for SC
software \citep{SmithEtAl2021}. Details can be found in
Section~\ref{ch_methods}.  Our methodology has been applied to MI software
\citet{Dong2021} and Lattice Boltzmann Solvers \citet{Michalski2021}.  This
methodology builds off prior work to assess the state of the practice for such
domains as Geographic Information Systems \citep{smith2018state}, Mesh
Generators \citep{smith2016state}, Seismology software
\citep{Smith2018Seismology}, and Statistical software for psychology
\citep{smith2018statistical}.  In keeping with the previous methodology, we have
maintained the constraint that the work load for measuring a given domain should
be feasible for a team as small as one person, and for a short time, ideally
around a person month of effort. A person month is considered to be $20$ working
days ($4$ weeks in a month, with $5$ days of work per week) at $8$ person hours
per day, or $20 \times 8 = 160$ person hours.

With our methodology, we first choose an SC domain (in the current case MI) and
identify a list of about 30 software packages. (For measuring MI we used 29
software packages.)  We approximately measure the qualities of each package by
filling in a grading template. Compared with our previous methodology, the new
methodology also includes repository based metrics, such as the number of files,
number of lines of code, percentage of issues that are closed, etc.  With the
quantitative data in the grading template, we rank the software with the
Analytic Hierarchy Process (AHP) (Details are found in
Section~\ref{ch_background}). After this, as another addition to our previous
methodology, we interview some of the development teams to further understand
the status of their development process.

@@

In this section, we introduce the relevant software categories (Section
\ref{sec_software_categories}). The categories will be used when collecting and
assessing the software packages.  We also review the definitions of software
qualities (Section \ref{sec_software_quality}) and provide an overview of 
AHP (Section \ref{sec_AHP}).

@@

Quality is defined as a measure of the excellence or worth of an entity.  As is
common practice, we do not think of quality as a single measure, but rather as a
set of measures.  That is, quality is a collection of different qualities, often
called ``ilities.''  Below we list the 10 qualities of interest for this study,
as summarized in \citet{SmithEtAl2021-QDef}.  The order of the qualities follows
the order used in \cite{GhezziEtAl2003}, which puts related qualities (like
correctness and reliability) together.  Moreover, the order is roughly the same
as the order qualities are considered in practice.  \wss{To be completed - not
currently in sync with QDef paper, QDef paper not completely done.  Look at LBM
paper too.}

@@

If there are $n$ alternatives, for criterion $i = 1, 2, ... , m$, we can create
an $n\times n$ matrix $B_i$ to record the relative preferences between these
choices for each criterion. The way of generating $B_i$ is similar to the one
for $A$. However, rather than comparing the importance between criteria, we
pairwise decide how much we favour one alternative over the other. We use the
same method to calculate the local priority vector for each $B_i$.  The local
priority vector in this case ranks the $n$ alternatives for criterion $i$.

@@

Domain experts may be recruited from academia or industry.  The only
requirements are knowledge of the domain and a willingness to be engaged in the
assessment process.  The Domain Expert does not have to be a software developer,
but they should be a user of domain software.  Given that the domain experts are
likely to be busy people, the measurement process cannot put to much of a burden
on their time.  For the current assessment, our Domain Expert (and paper
co-author) is Dr.\ Michael Noseworthy, Professor of Electrical and Computer
Engineering at McMaster University, Co-Director of the McMaster School of
Biomedical Engineering, and Director of Medical Imaging Physics and Engineering
at St. Joseph's Healthcare, Hamilton, Ontario, Canada.

@@

The repository measures help us in many ways. Firstly, they help us get a fast
and accurate project overview. For example, the number of commits over the last
12 months shows how active a project has been, and the number of stars and forks
may reveal its popularity. Secondly, the results may affect our decisions
regarding the grading scores for some software qualities. For example, if the
percentage of comment lines is low, we double-check the understandability of the
code; if the ratio of open versus closed pull requests is high, we pay more
attention to maintainability.

@@

\section{Comparison to Community Ranking} \label{Sec_VsCommunityRanking}

To address \rqref{RQ_CompareHQ2Popular}, Table~\ref{tab_ranking_vs_GitHub}
compares our LBM software package rankings against their popularity in the
research community, as estimated by repository stars and watches. We answer this
question by grading the qualities of the software, collecting ratings from the
MI community, then conducting two comparisons:
\begin{itemize}
\item comparing our ranking with the community ratings on GitHub, such as GitHub
stars, number of forks, and number of people watching the projects (Section~\ref{sec_grading_software});
\item comparing top-rated software designated by our methodology with the ones
recommended by our domain experts (Section \ref{sec_software_selection}).
\end{itemize}

\section{Comparison Between LBM and Research Software for Artifacts}
\label{Sec_CompareArtifacts}

As part of filling in the measurement template the LBM packages were examined
for the presence of artifacts, which were then categorized by frequency. We have
grouped them into common, less common, and rare artifacts in
Table~?. Common artifacts were found in 16 to 24 ($>$63\%)
of the software packages. Less common artifacts were found in 8 to 15 (30-63\%)
of the software packages. Rare artifacts were found in 1 to 7 ($<$30\%) of the
software packages.  In this section we answer \rqref{RQ_CompareArtifacts} by
comparing the artifacts that we observed in LBM repositories to those observed
and recommended for research software in general. Our comparison may point out
areas where some LBM software packages fall short of current best practices.
This is not intended to be a criticism of any existing packages, especially
since in practice not every project needs to achieve the highest possible
quality.  However, rather than delve into the nuances of which software can
justify compromising which practices we will write our comparison under the
ideal assumption that every project has sufficient resources to match best
practices.

\section{Comparison of Tool Usage Between LBM and Other Research Software}
\label{Sec_CompareTools}

Software tools are used to support the development, verification, maintenance,
and evolution of software, software processes, and artifacts \citep[p.\
501]{GhezziEtAl2003}. Many tools are used by LBM software packages, as
summarized in Table~?.  The tools are subdivided into development
tools, dependencies, and project management tools.  In this section we answer
\rqref{RQ_CompareToolsProjMngmnt} by comparing aspects of tool usage in LBM
software packages to their utilization in the research software community in
general.

\section{Comparison of Principles, Process and Methodologies to Research Software in General} \label{Sec_CompareMethodologies}

This section answers research question \rqref{RQ_CompareMethodologies} by
comparing the principles, processes and methodologies used for LBM software to
what can be gleaned from the literature on research software in general. In our
data collection for LBM software, the software development process is not
explicitly indicated in the artifacts for most of the packages. However, during
our interviews one developer (ESPResSo) told us their non-rigorous development
model is like a combination of agile and waterfall. Employing a loosely defined
process makes sense for LBM software, given that the teams are generally small
and self-contained.  Although eleven of the packages explicitly convey that they
would accept outside contributors, generally the teams are centralized, often
working at the same institution.  Working at the same institution means that an
informal process can show success, since informal conversations are relatively
easy to have.

@@

We identified the use of unit testing in less than half of the 29 projects. On
the other hand, the interviewees believed that testing (including usability
tests with users) was the top solution to improve correctness, usability, and
reproducibility. One pain point in the development process is the lack of access
to real-world datasets for testing. The developers' strategies to address this
are summarized in Section~\ref{ch_interview} \ad{to be fixed}. One threat to correctness is:
with huge datasets for testing, the tests are expensive and time-consuming.
Three interviewees endorsed self tests / automated tests, which may save time
for testing.

@@

@@

\item \textbf{Remove privacy from sensitive data.} For data with sensitive
information, a team can ask the data owner to remove such information or add
noise to protect privacy. One example is using de-identified copies of medical
images for testing.

@@

% Overview of scientific computing
We define Scientific Computing (SC) as ``the use of computer tools to analyze or
simulate mathematical models of real world systems of engineering or scientific
importance so that we can better understand and predict the system’s behaviour''
\citep{Smith2006}. Many researchers consider SC as the third pillar of science
and engineering, along with theory and experiment \citep{Landau2005}. Almost all
areas in science and engineering use computers for modelling \citep{Golub2014},
and software plays an essential role in modern scientific research
\citep{Hannay2009, Wilson2014}.  Software development in SC depends on three
fields of knowledge: engineering or scientific domain knowledge, mathematical
algorithm knowledge, and computational algorithm knowledge \citep{Landau2005,
Mehta2015}. Thus, most SC software developers are scientists in SC domains
\citep{Wilson2014}. However, they do not always use modern software development
techniques, tools, and methods \citep{Wilson2014}. Therefore, we developed a
methodology for assessing the state of the practice for SC software
\citep{SmithEtAl2021}. We apply this process to Medical Imaging (MI) software
that belongs to a specific domain of SC.
